---
title: "Investigating Cohort Extraction"
author: "Ted Laderas"
date: "5/4/2017"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(forcats)
library(knitr)
```

## Looking at the Cohort

Here I report on the various statistics of the cohort so we can decide whether our risk groups are realistic or not.

```{r cars}
cohort <- read.csv("~/Code/OHDSIqueries/data/cohorts_05022017.csv")
#recode these variables as categorical
cohort$outcome <- factor(cohort$outcome)
cohort$htn <- factor(cohort$htn)
cohort$htn_treatment <- factor(cohort$htn_treatment)
cohort$diabetes <- factor(cohort$diabetes)
#add age
cohort <- cohort %>% mutate(age=2017-year_of_birth)

```

Here are the overall summary statistics of our cohort:

```{r}
returnReport <- function(cohort, totalSize){
  cohortOut <- cohort %>% select(age, race, outcome, diabetes,
                                 htn, htn_treatment, bmi,
                                 tchol)
  fct_count(cohort$outcome)
  outcomeTable <- table(cohort$outcome)
  percentOfTotal <- signif(nrow(cohort)/totalSize*100, digits=3)
  incidence <- signif(outcomeTable[2]/outcomeTable[1]*100,digits = 3)
  print(paste0("Number in subgroup: ", nrow(cohort), 
               " (", percentOfTotal, "%)"))
  print(paste("Incidence of CVD in subgroup:", incidence, "%"))
  print(summary(cohortOut))
  return(list(percent=percentOfTotal, incidence=incidence))
  }


returnReport(cohort, nrow(cohort))

condHTNtreat <- table(cohort$htn, cohort$htn_treatment)/nrow(cohort)
out <- cut(cohort$age, breaks=c(0,20, 40, 55, 70,90))
cohort <- data.frame(cohort, ageCategory=out)

#cohort <- cohort %>% mutate(ageCategory = ntile(age,n=4))

tabOut <- table(out)
ageProb <- tabOut/sum(tabOut)

condHTNAge <- table(cohort$ageCategory, cohort$htn)

condHTNAgeprob <- condHTN/sum(condHTN)
```
These are the subgroups that we have defined in the dataset.

```{r}
subGroups <- read.csv("~/Code/OHDSIqueries/scripts/subgroups.csv", stringsAsFactors = FALSE)

knitr::kable(subGroups)

outList <- lapply(subGroups$defintion, function(x){return(strsplit(x, ",")[[1]])})
names(outList) <- subGroups$cohortNumber
```

Here I return the first five subgroups and their statistics. Note that I don't calculate the 6-10th subgroups because we don't yet have smoking status.

Overall, we have very few total cholesterol numbers. 

```{r}
cohortList <- lapply(names(outList)[1:5], function(x){
  out <- cohort %>% filter_(.dots=outList[[x]])
  print(paste("cohort number:", x))
  print(paste(outList[[x]], collapse = ", "))
  out <-returnReport(out, totalSize = nrow(cohort))
  out <- data.frame(subGroups[x,],out)
})
```

Here are the groups and their percent of the total data and their CVD incidence.

```{r}

kable(do.call(rbind, cohortList))
