---
title: "SNP generation"
author: "Ted Laderas"
date: "March 22, 2017"
output: html_document
bibliography: ~/Dropbox/zotero/MyLibrary.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SNP Generation

Here I explore a number of SNP approaches for case/control studies. The first is a resampling approach based on HapMap

http://www.allelefrequencies.net

```{r}
library(hapmapsnp6)
```

## SNPs of Interest

[SNPedia](http://snpedia.com/index.php/Heart_disease) lists a number of potentially interesting SNPs to model:

[Rs4665058](http://snpedia.com/index.php/Rs4665058), 4x risk if have two copies of the SNP (A;A), 2 X (A;C) and (C;C) - wild type. Chr 2, 159333698

[Rs283206](http://snpedia.com/index.php/Rs2383206), Smaller risk, (Chr 9, 22115027). 1.7 x risk (G;G), 1.4x risk (A;G) and (A;A) wild type. Co-morbidity: diabetes

[Rs10757278](http://snpedia.com/index.php/Rs10757278) (Chr 9, 22124478). (A;A) 0.78 x risk of heart disease, (A;G) 1.3 x risk, (G;G) 1.6x risk for heart disease. Co-morbidity: diabetes

[Rs1333049](http://snpedia.com/index.php/Rs1333049) (Chr 9, 22125504). (G;G) Normal (WT); (C;G) 1.5 increased risk; (C;C) 1.9x increased risk.

## Simulation Approach

[SimuPOP](http://simupop.sourceforge.net) is a python package that seems to have most of the simulation functionality that we need. Specifically, the [GWAS example](https://github.com/BoPeng/simuPOP-examples) [@peng_forward-time_2010] seems to be a good case to base it on.

Base prevalence of CVD: According to [Heart.org](https://www.heart.org/idc/groups/heart-public/@wcm/@sop/@smd/documents/downloadable/ucm_449846.pdf)

20-30 - prevalence is 0.8% for CVD.